Applicability of SSM and UML for Designing a Search Application for the British Broadcasting Corporation (BBC)

### 2014-11-05 academy-talk.tex
Final draft for delivering.

% Background
\begin{frame}
  \frametitle{About Me}
  \begin{itemize}
    \item Senior Software Engineer
    \item Future Media: Content Discovery: Search
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Design Goals}
  \framesubtitle{Requirements}
  \begin{itemize}
    \item Existing search application serves a wide variety of use cases
    \item The Search application should integrate where possible with other existing systems
    \item Website is very high traffic
    \item Different parts of the overall website are maintained by distinct teams
    \item Varying applications, data stores and content management systems
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{What is SSM?}
  \framesubtitle{Peter Checkland}
  \begin{center}
    \includegraphics[height=0.8\textheight]{ssma.jpg}
    \quad
    \includegraphics[height=0.8\textheight]{stsp.jpg}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Problem Contextualisation through SSM}
  \begin{itemize}
    \item Claim: ``Hard systems'' (e.g. UML) take an ontological view:
    \begin{itemize}
      \item What components are there?
      \item Who are the actors?
      \item What are the interactions between actors and components?
      \item What are the computations or workflows steps/stages?
    \end{itemize}
    \pause \item Claim: SSM takes an epistemological view:
    \begin{itemize}
      \item How does the holistic search system transform user needs to content and information?
      \item How can the purposeful activity be monitored?
      \item How is the activity observed by different observers?
      \item How can different observers inform the activity?
    \end{itemize}
    \pause \item Dogan and Henshaw (2010) showed a transtion from soft systems to hard models
  \end{itemize}
\end{frame}

@@

% Designing behaviour for indirect users
\begin{frame}
  \frametitle{Use Case Extraction}
  \framesubtitle{Meta-level Users}
  \begin{itemize}
  \item Meta-level users do not transition to the use case models directly
  \item We should still consider how meta-level users are affected by emergent properties of the holistic system
    \begin{itemize}
    \item Actors
    \item Journalists
    \item Parents
    \item Other Future Media products?
    \item Future Media staff
    \end{itemize}
  \end{itemize}
\end{frame}

% Designing behaviour for direct users
\begin{frame}
  \frametitle{Use Case Extraction}
  \framesubtitle{Direct Users}
  \begin{itemize}
    \item Direct users transition directly to actors into a UML use case diagram
    \item Use cases should derive from and be justifiable by the rich picture
    \item Direct users divide up into \emph{content production} and \emph{audience}
    \item Examples of each:
    \begin{itemize}
      \item News editors
      \item Users of the search web application
    \end{itemize}
  \end{itemize}
\end{frame}

@@

% Discussion and Analysis -- Domain Model -- Polymorphism
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Domain Model}
  \begin{itemize}
    \item Large, maximal data model useful as communication artefact
    \item Aggregate models hard to keep in sync with changes in respective domains
    \item Does Search need to model all of this?
      \begin{itemize}
        \item Polymorphic supertype -- ``Thing'' class?
        \item Facade -- ``SearchResult'' class?
      \end{itemize}
  \end{itemize}
\end{frame}

% Discussion and Analysis -- Domain Model -- Duck Typing
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Schemalessness and Duck Typing}
  \begin{itemize}
    \item Statically typed approach: A \emph{Place} always has a ``today's weather forecast'' attribute
    \item Dynamically (``duck'') typed approach: any entity can have a ``today's weather forecast'' attribute and we display this forecast if that attribute is present
    \item Allows other entities in future to have forecasts too (e.g. \emph{Events} such as football matches)
    \item Search application can store all fields from upstream sources and choose when to start using them
  \end{itemize}
\end{frame}

% Discussion -- More SSM: 3 Es?
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Overall Model: Use cases, behaviour and domain}
  \begin{itemize}
    \item Looks good
    \item Agile? MVP? Test assumptions?
    \item How do we monitor for ``3 Es''?
      \pause \begin{itemize}
        \item Efficacy -- Does the system do what we intended?
        \item Efficiency -- Does it minimise use of resources?
        \item Effectiveness -- Does it fit into our long term goals?
      \end{itemize}
  \end{itemize}
\end{frame}

% Shortcomings -- Agile?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Detailed Design and Agile}
  \begin{itemize}
    \item High-level design lacks detail (not surprisingly)
    \item Agile encourages delivering something tangible early
    \item Shouldn't the full detailed design emerge iteratively?
    \item Can we iteratively build up and maintain hard models?
    \item Can we defer pending questions in our hard models until we know more?
    \item SSM encourages continuous feedback cycles -- does this align with Agile?
    \item Bustard, Dave, and Frank Keenan. ``Soft Systems Methodology: An Aid to Agile Development?.''
      In \emph{Information Systems Development}, pp. 25-38. Springer US, 2009.
  \end{itemize}
\end{frame}

>>>>

% Background
\begin{frame}
  \frametitle{About Me}
  \begin{itemize}
    \item Senior Software Engineer
    \item Future Media: Content Discovery: Search
    \item Presented at EMMSAD workshop as part of CAiSE 2014
      \begin{itemize}
        \item Exploring Modelling Methods for Systems Analysis and Design
        \item 26th International Conference on Advanced Information Systems Engineering
      \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Design Goals}
  \framesubtitle{Requirements}
  \begin{itemize}
    \item Existing search application serves a wide variety of use cases
    \item The Search application should integrate where possible with other existing systems
    \item BBC website is very high traffic
    \item Different parts of the BBC website are maintained by distinct teams
    \item Varying applications, data stores and content management systems
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{What is SSM?}
  \framesubtitle{Framework for ``messy'' problems}
  \begin{center}
    \includegraphics[height=0.8\textheight]{ssma.jpg}
    \quad
    \includegraphics[height=0.8\textheight]{stsp.jpg}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Problem Contextualisation through SSM}
  \begin{itemize}
    \pause \item Claim: ``Hard systems'' (e.g. UML) take an ontological view:
    \begin{itemize}
      \item What components are there?
      \item Who are the actors?
      \item What are the interactions between actors and components?
      \item What are the computations or workflows steps/stages?
    \end{itemize}
    \pause \item Claim: SSM takes an epistemological view:
    \begin{itemize}
      \item How does the holistic search system transform user needs to content and information?
      \item How can the purposeful activity be monitored?
      \item How is the activity observed by different observers?
      \item How can different observers inform the activity?
    \end{itemize}
    \pause \item Dogan and Henshaw (2010) showed a transtion from soft systems to hard models
  \end{itemize}
\end{frame}

@@

% Designing behaviour for indirect users
\begin{frame}
  \frametitle{Use Case Extraction}
  \framesubtitle{Meta-level Users}
  \begin{itemize}
  \item Meta-level users do not transition to use cases directly
  \item We should still consider how meta-level users are affected by emergent properties of the holistic system
    \begin{itemize}
    \item Actors
    \item Journalists
    \item Parents
    \item Other Future Media products?
    \item Future Media staff
    \end{itemize}
  \end{itemize}
\end{frame}

% Designing behaviour for direct users
\begin{frame}
  \frametitle{Use Case Extraction}
  \framesubtitle{Direct Users}
  \begin{itemize}
    \item Direct users transition directly to actors in a UML use case diagram
    \item Use cases should derive from and be justifiable by the rich picture
    \item Direct users divide up into \emph{content production} and \emph{audience}
    \item Examples of each:
    \begin{itemize}
      \item News editors
      \item Users of the search web application
    \end{itemize}
  \end{itemize}
\end{frame}

@@

% Discussion and Analysis -- Domain Model -- Polymorphism
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Domain Model}
  \begin{itemize}
    \item Large, maximal data model useful as communication artefact
    \item Aggregate models hard to keep in sync with changes in respective domains
    \item Does Search need to model all of this?
      \begin{itemize}
        \item Supertype -- ``Thing'' class?
        \item Facade -- ``SearchResult'' class?
      \end{itemize}
  \end{itemize}
\end{frame}

% Discussion and Analysis -- Domain Model -- Duck Typing
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Schemalessness and Duck Typing}
  \begin{itemize}
    \item \url{http://martinfowler.com/articles/schemaless/}
    \item Statically typed approach: A \emph{Place} always has a ``today's weather forecast'' attribute
    \item Dynamically (``duck'') typed approach: any entity can have a ``today's weather forecast'' attribute and we display this forecast if that attribute is present
    \item Allows other entities in future to have forecasts too (e.g. \emph{Events} such as football matches)
    \item Search application can store all fields from upstream sources and choose when to start using them
    \item Contextual validation enforces ``required'' fields are always populated
  \end{itemize}
\end{frame}

% Discussion -- More SSM: 3 Es?
\begin{frame}
  \frametitle{Discussion}
  \framesubtitle{Overall Model: Use cases, behaviour and domain}
  \begin{itemize}
    \item Looks good\ldots at a high level
    \item Agile? MVP? Test assumptions?
    \item How do we monitor for ``3 Es''?
      \pause \begin{itemize}
        \item Efficacy -- Does the system do what we intended?
        \item Efficiency -- Does it minimise use of resources?
        \item Effectiveness -- Does it fit into our long term goals?
      \end{itemize}
  \end{itemize}
\end{frame}

% Shortcomings -- Agile?
\begin{frame}
  \frametitle{Evaluation and Difficulties}
  \framesubtitle{Detailed Design and Agile: Using SSM as part of BBC development practices?}
  \begin{itemize}
    \item High-level design lacks detail (not surprisingly)
    \item Agile encourages delivering something tangible early
    \item Shouldn't the full detailed design emerge iteratively?
    \item Can we iteratively build up and maintain hard models?
    \item Can we defer questions in our hard models until we know more?
    \item SSM encourages continuous feedback cycles -- Agile?
    \item BDD encourages adding business context to features
    \item Bustard, Dave, and Frank Keenan. ``Soft Systems Methodology: An Aid to Agile Development?.''
      In \emph{Information Systems Development}, pp. 25-38. Springer US, 2009.
  \end{itemize}
\end{frame}
